\section{Conclusion}

We presented a multi-agent framework decomposing human user simulation into specialized components: a User Agent for orchestration, State Tracking Agent for task management, and Message Attributes Generation Agent for behavioral control. Through persona control and task state grounding, our framework enables realistic simulation in conversational domains. Validating it in restaurant guest ordering scenarios demonstrates 102.6\% improvement in simulation quality over single-LLM baselines, with significant gains in persona adherence (19.9\%), behavioral variance (284.5\%), and task completion accuracy (29.1\%). These results validate that decomposing human simulation into specialized agents with coordinated state management yields superior performance across simulation quality dimensions, establishing our multi-agent architecture as effective for realistic human simulation in interactive AI systems.

\paragraph{Limitations and Ethical Considerations}
The multi-agent architecture incurs substantial computational costs (124\% more tokens, 356\% higher latency), limiting resource-constrained deployment. Each domain requires specialized prompt engineering and state design. Behavioral modeling lacks complex human behaviors like indecision, social dynamics, or cultural nuances~\cite{Wang2025SurveyLLMAgents}. Validation is English-only, single-domain with 60 test cases, restricting cross-cultural, multilingual, and real-world insights. Systems must maintain transparency and avoid bias perpetuation through auditing~\cite{Li2025LLMGeneratedPersona}. While democratizing conversational AI testing, the framework shouldn't impersonate individuals without consent or manipulate users believing they interact with real humans.

\paragraph{Future Work and Applications}
Key directions include adaptive persona evolution enabling dynamic behavioral adjustment~\cite{Cheng2024AutoPal}, multi-modal integration for voice-based prosodic features, visual gesture recognition, and emotional sentiment tracking~\cite{Chu2024MultimodalEmotional}, and cross-domain validation across customer support, e-commerce ordering~\cite{Xiang2024TransformerEcommerce}, healthcare consultations~\cite{Yu2024AIPatient}, educational tutoring~\cite{Maity2024GenerativeAI_ITS}, financial advisory, travel booking, and technical troubleshooting. Efficiency optimization through agent caching, selective tool invocation, and smaller specialized models could address computational overhead. The core principle of decomposing human behavioral simulation into specialized agents managing persona consistency and task state provides a foundation for realistic user simulations across interactive domains. Our framework offers systematic high-quality interaction data generation for testing, evaluation, and quality assurance wherever validating human-system interactions is critical for system reliability and user experience.


